---
title: "River EDM"
author: "Melissa DeSiervo"
date: "4/11/2022"
output: html_document
---



Install necessary packages
```{r Upload packages}
library(dplyr)
library(tidyr)
library(lme4)
library(ggplot2)
library(here)
#devtools::install_github("ha0ye/rEDM")

library(rEDM)

```


annualstreams <- read.csv(here("data_working/across_sites_model_data.csv"), header=T)

autodailystreams <- read.csv(here("data_working/autotrophic_siteyears_daily.csv"), header=T)

autodailystreams$date<-as.Date(dailystreams$date)
autodailystreams<- dailystreams %>% mutate(PRannual=GPP_annual/ER_annual) 


dailyall_notclean <-readRDS(here("data_356rivers/high_quality_daily_metabolism_with_SP_covariates.rds"))



## see which streams are the most heterotrophic and autotrophic ####


##pull out stream nwis_02266200 nwis_01608500 and 2014-2016 ####


dailyall_subset <-subset(dailyall_notclean, site_name=="nwis_02266200" |site_name=="nwis_01608500"  )

annualstreamsubset_2014<-subset(dailyall_subset, year ==2014 )


###


##pull out stream nwis_13173600 (Snake river) ####


snakeriver_subset <-subset(dailyall_notclean, site_name=="nwis_13173600")



###


```{r plot ER and GPP over time for 2 streams in 2014}


ann_text <- data.frame(date=(20+min(annualstreamsubset_2014$date)), GPP=-20, label=c("Annual PR: 1.33","Annual PR: 0.025"), long_name=c("SOUTH BRANCH POTOMAC RIVER NEAR SPRINGFIELD, WV", "WHITTENHORSE CREEK NEAR VINELAND, FL"))


ERGPPtwostreams<-ggplot(data=annualstreamsubset_2014) +   geom_line(aes(x = date, y = GPP), color="chartreuse4", size=1.25)+  geom_line(aes(x = date, y = ER), color = "brown", size=1.25)+ylim(-25, 15)+labs(y=expression('GPP and ER (g '*~O[2]~ m^-2~d^-1*')'), x="")+ facet_wrap(~ long_name, ncol = 1)+theme_bw()+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ geom_text(data = ann_text,  mapping = aes(x = date, y = GPP, label = label), color=c("chartreuse4", "brown"))+geom_hline(yintercept=0,linetype=2)


```


```{r plot snake river in 2010}


ann_text2 <- data.frame(date=(20+min(snakeriver_subset$date)), GPP=-30, label=c("Annual PR: 1.6"))

  
ERGPPsnake<-ggplot(data=snakeriver_subset) +   geom_line(aes(x = date, y = GPP_filled), color="chartreuse4", size=1.25)+  geom_line(aes(x = date, y = ER_filled), color = "brown", size=1.25)+labs(y=expression('GPP and ER (g '*~O[2]~ m^-2~d^-1*')'), x="")+theme_bw()+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+geom_hline(yintercept=0,linetype=2)+ggtitle("Snake river")+ geom_text(data = ann_text2,  mapping = aes(x = date, y = GPP, label = label), color=c("chartreuse4"))


```





